---
title: "Variation over time"
author: "Laurent Berder"
output:
  html_document: default
  html_notebook: default
---

## Preparation
Load packages:
```{r libraries, message=FALSE, warning=FALSE}
library(data.table)
library(sqldf)
library(dplyr)
library(tidyr)
library(xlsx)
library(plotly)
```

Import data:
```{r import}
names <- read.csv("../input/NationalNames.csv")
```

The first thing I'll do to avoid having to use filters all through the code is to separate boys' names from girls' names:
```{r separate}
boys<-filter(names,Gender=="M")
girls<-filter(names,Gender=="F")
```
By the way, I realize that *nrow(boys)+nrow(girls)* does not equal *nrow(names)*: most of the 1880 data is missing its gender type.


## Sum of names and number of distinct names
I'm wondering if the number of names is evolving the same way as the number of data over the years.

First off, I'm summarizing data by year, showing the number of distinct names, as well as the total of names:
```{r count sum}
boycount<-group_by(boys,Year) %>% summarise(nbnames=n_distinct(Name),sum=sum(Count))
girlcount<-group_by(girls,Year) %>% summarise(nbnames=n_distinct(Name),sum=sum(Count))
```

I'm then using the package *Plotly* to create an interactive plot:
```{r plot sum}
ay <- list(
  tickfont = list(color = "orange"),
  overlaying = "y",
  side = "right"
)

p<-plot_ly(data=boycount, x=Year, y=nbnames, type = "scatter", mode="lines", name="Number of distinct boy names") %>%
  add_trace(data=boycount, x=Year, y=sum, type = "scatter", mode="lines", name="Sum of boy names", yaxis = "y2") %>%
  add_trace(data=girlcount, x=Year, y=nbnames, type = "scatter", mode="lines", name="Number of distinct girl names") %>%
  add_trace(data=girlcount, x=Year, y=sum, type = "scatter", mode="lines", name="Sum of girl names", yaxis = "y2") %>%
  layout(title="Evolution of number of names count and sum", yaxis2=ay)
p
```
It seems the number of distinct names has grown more rapidly over the past 60 years than the number of total names, bringing more variety.

This is even truer for girls' names, which historically have always had more diversity.

## Most common names
I'm then looking at the most commonly used names per year, and what percentage they represent.

First I need to summarize the data (I'm using the total sum which I've already calculated above):
```{r common names, warning=FALSE, message=FALSE}
commonboy<-sqldf("select year, max(Count), Name from boys group by Year") %>% left_join(boycount, by="Year")
colnames(commonboy)<-c("Year","Maximum","Name","nb","Total")
commonboy<-mutate(commonboy,Percent=Maximum/Total*100) %>% select(c(Year,Name,Percent))

commongirl<-sqldf("select year, max(Count), Name from girls group by Year") %>% left_join(girlcount, by="Year")
colnames(commongirl)<-c("Year","Maximum","Name","nb","Total")
commongirl<-mutate(commongirl,Percent=Maximum/Total*100) %>% select(c(Year,Name,Percent))
```

Then I'm plotting these figures with plotly again:
```{r common plot}
b <- list(family = "Courier New, monospace",
  size = 10,
  color = "#00008B")
g <- list(family = "Courier New, monospace",
  size = 10,
  color = "#FF4500", orientation="v")

p<-plot_ly(data=commonboy, x = Year, y = Percent, text = Name, mode = "markers+text+lines", textfont = b, textposition = "top middle", name="Most common boy's name") %>%
    add_trace(data=commongirl, x = Year, y = Percent, text = Name, mode = "markers+text+lines", textfont = g, textposition = "top middle", name="Most common girl's name") %>%
    layout(title="Most common names in percentage per gender and per year", showlegend=FALSE)
p
```
There may be work to do for a more readable result.

It's interesting to see that the classical couple **John - Mary** ruled from 1880 to 1922, with a gradually eroding preponderance (Mary remaining unbeated lead girl's name until 1946)

We can clearly see that the diversity increases, as the most common names takes less and less part in the composition of the global population.

The girls' curve is less regular than the boys', with a huge peak at the introduction of **Linda** in 1947-48, and another one only after the introduction of **Jennifer** in 1970.